Forecasting Gamma-ray Bursts using Gravitational Waves

### 2018-03-23 Paper/Manuscript.tex
mods

To represent our model three-IFO network, we introduce % the norm of the frequency-domain strain (\ref{eq:strain_FD}) 
%%
%
\be
\tilde{H}_i(f)\equiv \tilde{A}_i\, h_0\, f^{-7/6} \label{eq:Hi_s}
\ee
%
%%
for each IFO labelled by $i$, where
%%
%
\begin{align}
%\tilde{A}_i &\equiv A \langle |Q|^2\rangle^{1/2}_i=\f{1}{2}\sqrt{\f{5}{6}}\, \pi^{-2/3} \left\langle\, |Q_i(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\quad\text{and} \label{eq:A_tilde}\\
h_0 &= \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} \label{eq:h0}\, .
\end{align}
%
and $\tilde{A}_i =A\langle|Q|^2\rangle^{1/2}_i $ are given in Table~\ref{table:network_params}. %for $i=1,2,3$. %Clearly, the strain and thus the SNR at each of our imaginary interferometers scales as $\langle |Q|^2\rangle^{1/2}_i$, thus
%%
The SNR for each interferometer can now be written as
%%
%
\begin{align}
\rho_i &= 2\tilde{A}_i\, h_0 \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{n}(f)} \right]^{1/2}\label{eq:SNR_i}.
\end{align}
%
We arbitrarily make the following assignments for our network: $h=1, m=2, l=3$ representing high, medium, low strain sensitivity, respectively. 
We summarize our network in the following table.
%%

\begin{table}[h]
\centering
\begin{tabular}{c|c|c|c}
 IFO$_i$ & configuration %& $\langle |Q_i|^2\rangle$ 
 & $\tilde{A}_i$ & $\tilde{H}_i(f)$\\
 %\hline
 \hline
 1 & high ($h$) &  $\tfrac{4}{5}A$ & $\tfrac{4}{5}A h_0\, f^{-7/6}$ \T\\
 2 & medium ($m$) & $\tfrac{2}{5}A$ & $ \tfrac{2}{5}A h_0 \, f^{-7/6}$\B\T \\
 3 & low ($l$)  & $\left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A$ & $ \left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A h_0 \, f^{-7/6}$\B \\
 \hline
\end{tabular}
\caption{%The averaged quality factors $\langle|Q|^2\rangle$ of 
Our model for the Advanced-LIGO-Virgo network ca. 2020. %of three L-shaped interferometers (IFOs). 
Column four lists the magnitude of the frequency-domain strain at each IFO via Eq.~(\ref{eq:Hi_s}).
``Configuration'' describes how ideally oriented/positioned an IFO is with respect to a compact binary source. 
High is $2\sigma$ above the average as explained above.
Medium is the RMS-averaged (one $\sigma$) response. Recall $A= \pi^{-2/3} \sqrt{5/24}$. 
%Note that for an interferometer where the angle between the arms is $\alpha$, $A$ rescales to $A\sin\alpha$.
}\label{table:network_params}
\end{table}
%
%
We can now write the total SNR for our network
%%
%
\begin{align}
\rho_\text{tot} &= \sqrt{\rho^2_h+\rho^2_m+\rho^2_l} \label{eq:SNR_total_3IFOs} 
%	      &= 2A h_0 \sqrt{1+\tfrac{2}{5}+\tfrac{4}{25}} \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2}\nn\\
%	      &= \sqrt{\frac{13}{10}}\,\pi^{-2/3}\,\f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6}\left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2} \label{eq:SNR_total_3IFOs_v2}.
\end{align}
%
%%
with $\rho_h, \rho_m,\rho_l$ given by Eq.~(\ref{eq:SNR_i}) and Table~\ref{table:network_params}.
%Although we made several simplifying assumptions to obtain $\rho_\text{tot}$, these should not affect our total SNR estimation too much.
%For example, averaging over all the angles underestimates the SNR of an interferometer with a near-ideal orientation with respect to the incoming GWs
%while also overestimating the SNR of a badly oriented detector. Rewriting Eq.~(\ref{eq:SNR_total}) as Eq.~(\ref{eq:SNR_total_3IFOs}) compensates
%for these.
%
%In Sec.~\ref{sec:GW170817} we test our model network with GW170817 whose signal got picked up by the LIGO detectors around $f=30$\,Hz.
%To do so we will change the lower limit of the integral in Eq.~(\ref{eq:SNR_total_3IFOs}) from $0$ to $30$.
%For our advance warning times we will further use separate initial frequencies for each inteferometer as we will explain in Sec.~\ref{sec:results}.

\subsection{Einstein Telescope configuration: triangular topology}\label{sec:ET_topo}
Current design of Einstein Telescope is based on a [equilateral] triangular configuration with 10\,km armlengths \cite{ET_doc}.
Within this equilateral triangle, ET will consist of three V-shaped cryogenic interferometers housed underground to
sigificantly reduce seismic and gravity-gradient noises \cite{ET_doc}.
This closed topology will allow the detector to form a null stream completely devoid of a signal \cite{Sathyaprakash:2012jk},
which can be used to rule out spurious events \cite{Wen:2005ui}.
As we will show below, the response function and the quality factor of the triangular configuration are independent of the azimuthal angle $\phi$
hence ET will have no blind spots \cite{Regimbau:2012ir}.
However, the $60^\circ$ arm separation reduces the strain of a V-shaped interferometer by a factor of $\sin 60^\circ=\sqrt{3}/2$ compared
to an L-shaped interferometer. So for a single V-IFO the antenna patterns of (\ref{eq:F_plus}, \ref{eq:F_cross}) become
%
\be
F^1_+(\theta,\phi,\psi)= \f{\sqrt{3}}{2} F_+(\theta,\phi,\psi), \qquad F^1_\times(\theta,\phi,\psi)= \f{\sqrt{3}}{2} F_\times(\theta,\phi,\psi)\label{eq:F1plus_F1cross}\, ,
\ee
%
where the superscript $1$ labels one of the three V's of the triangle.
Since the other two V IFOs lie in the same plane as the first one, their antenna patterns can be obtained simply by rotating in the azimuthal
direction by $120^\circ, 240^\circ$, respectively
%
\begin{align}
 F^2_{+,\times}(\theta,\phi,\psi) &= F^1_{+,\times}(\theta,\phi+2\pi/3,\psi), \label{eq:F2plus_F2cross}\\
 F^3_{+,\times}(\theta,\phi,\psi) &= F^1_{+,\times}(\theta,\phi-2\pi/3,\psi) \label{eq:F3plus_F3cross}.
\end{align}
%
The invidual antenna responses combine to cancel the $\phi$ and $\psi$ dependence of the combined power pattern
%
\be
F^2_\bigtriangleup \equiv \sum_{A=1}^3 \left(F^A_+\right)^2+\left(F^A_\times\right)^2 = \f{9}{32}\left(1+6\cos^2\theta + \cos^4\theta\right). \label{eq:ET_power_pattern}
\ee
%
From this, we immediately have that the RMS average is given by $\langle F_\bigtriangleup\rangle =3/\sqrt{10}$ and the minium value is $F_\bigtriangleup(\pi/2)=3/\sqrt{32}$;
%As $F^2 > 0$ for all $\theta$ with its RMS average $\sqrt{2/5}\simeq 63.2\%$ of optimal $F$ and its
thus a triangular detector made-up of three V-IFOs has all-sky coverage.

For the total SNR accumulated in a triangular detector we have
%
\begin{align}
\rho^2_{\text{tot},\bigtriangleup} & = \sum_{A=1}^3 \rho_A^2 \nn \\
 & =  \f{5}{6}\, \pi^{-4/3} \f{c^2}{D^2}\left(\f{G M_c}{c^3}\right)^{5/3}|Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 \int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \, ,\label{eq:ET_SNRsq}
\end{align}
where
%
\be
|Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 = \sum_{A=1}^3 \left(F^A_+\right)^2\f{1+\cos^2\iota}{2}  + \left(F^A_\times\right)^2 \cos\iota \label{eq:Q_sq_ET}
\ee
%
which is independent of $\phi$ \cite{Regimbau:2012ir}. We once again take the RMS average of this quantity
%$Q(\theta,\phi,\psi,\iota)$ is the quality factor of the detector.
%
\be
\bar{Q}^2_\bigtriangleup\equiv\langle |Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 \rangle_{\theta,\phi,\psi,\iota} = \f{9}{25}\label{eq:RMS_ET}
\ee
%
which is $5/2$ times higher than a single L-shaped IFO's RMS average of 4/25 [cf. Eq.~(\ref{eq:RMS_th_ph_ps_iota})].

As we can not know the source sky position or orientation we will use the RMS-averaged SNR for our computations %(\ref{eq:ET_SNRsq})
%
\be
\rho_{\text{tot},\bigtriangleup} = \f{36}{25}A^2 h_0^2 \int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \label{eq:ET_SNR},
\ee
where $h_0$ is given by Eq.~(\ref{eq:h0}) and $A=\pi^{-2/3}\sqrt{5/24}$ as before. % from Sec.\ref{sec:LIGO_topo}. 
Accordingly, we define an RMS-averaged frequency-domain
strain
%
\begin{align}
\tilde{H}_\bigtriangleup(f, D)&\equiv A\,  \bar{Q}_\bigtriangleup\, h_0 f^{-7/6} \nn \\
   & = \f{1}{2}\sqrt{\f{3}{10}}\, \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} f^{-7/6}. \label{eq:H_delta}
\end{align}
%
It is this strain that we will use as a typical ET source; thus in this sense,
$\tilde{H}_\bigtriangleup(f,D)$ is like the $\tilde{H}_i(f)$ of Table~\ref{table:network_params}.
But whereas in the case of a three L-IFO network we represented the detector response due to a single source in terms of three different
values $\tilde{H}_i(f)$, for ET we will only consider $\tilde{H}_\bigtriangleup(f)$ as we have already angle-averaged over all three V-IFOs antenna patterns.
%$\tilde{H}_\bigtriangleup(f)$, as given by Eq.~(\ref{eq:H_delta}), depends only on the luminosity distance and the chirp mass.
In Sec.~\ref{Sec:ETB} we will consider BNS systems inspiralling at varying luminosity distances as typical sources for ET.
To compute the resulting SNRs, we will slightly modify Eq.~(\ref{eq:ET_SNR}) and use $\sqrt{S_n(f)}$ for the B and C configurations of ET.

Before we present our results, we list the various idealizations/simplifications that we adopt to simplify our computations.

>>>>

To represent our model three-IFO network, we introduce the norm of the frequency-domain strain (\ref{eq:strain_FD}) 
%%
%
\be
\tilde{H}_i(f)\equiv \tilde{A}_i\, h_0\, f^{-7/6} \label{eq:Hi_s}
\ee
%
%%
for each IFO labelled by $i$, where
%%
%
\begin{align}
%\tilde{A}_i &\equiv A \langle |Q|^2\rangle^{1/2}_i=\f{1}{2}\sqrt{\f{5}{6}}\, \pi^{-2/3} \left\langle\, |Q_i(\theta,\phi,\psi,\iota)|^2\,\right\rangle^{1/2}\quad\text{and} \label{eq:A_tilde}\\
h_0 &= \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} \label{eq:h0}\, .
\end{align}
%
and $\tilde{A}_i =A\langle|Q|^2\rangle^{1/2}_i $ are given in Table~\ref{table:network_params}. %for $i=1,2,3$. %Clearly, the strain and thus the SNR at each of our imaginary interferometers scales as $\langle |Q|^2\rangle^{1/2}_i$, thus
%%
The SNR for each interferometer can now be written as
%%
%
\begin{align}
\rho_i &= 2\tilde{A}_i\, h_0 \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{n}(f)} \right]^{1/2}\label{eq:SNR_i}.
\end{align}
%
We arbitrarily make the following assignments for our network: $h=1, m=2, l=3$ representing high, medium, low strain sensitivity, respectively. 
We summarize our network in the following table.
%%

\begin{table}[h]
\centering
\begin{tabular}{c|c|c|c}
 IFO$_i$ & configuration %& $\langle |Q_i|^2\rangle$ 
 & $\tilde{A}_i$ & $\tilde{H}_i(f)$\\
 %\hline
 \hline
 1 & high ($h$) &  $\tfrac{4}{5}A$ & $\tfrac{4}{5}A h_0\, f^{-7/6}$ \T\\
 2 & medium ($m$) & $\tfrac{2}{5}A$ & $ \tfrac{2}{5}A h_0 \, f^{-7/6}$\B\T \\
 3 & low ($l$)  & $\left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A$ & $ \left(\tfrac{4}{5}-\sqrt{\tfrac{2}{5}}\right)A h_0 \, f^{-7/6}$\B \\
 \hline
\end{tabular}
\caption{%The averaged quality factors $\langle|Q|^2\rangle$ of 
Our model for the Advanced-LIGO-Virgo network ca. 2020. %of three L-shaped interferometers (IFOs). 
Column four lists the magnitude of the frequency-domain strain at each IFO via Eq.~(\ref{eq:Hi_s}).
``Configuration'' describes how ideally oriented/positioned an IFO is with respect to a compact binary source. 
High is $2\sigma$ above the average as explained above.
Medium is the RMS-averaged (one $\sigma$) response. Recall $A= \pi^{-2/3} \sqrt{5/24}$. 
%Note that for an interferometer where the angle between the arms is $\alpha$, $A$ rescales to $A\sin\alpha$.
}\label{table:network_params}
\end{table}
%
%
We can now write the total SNR for our three L-IFO network
%%
%
\begin{align}
\rho_\text{tot} &= \sqrt{\rho^2_h+\rho^2_m+\rho^2_l} \label{eq:SNR_total_3IFOs} 
%	      &= 2A h_0 \sqrt{1+\tfrac{2}{5}+\tfrac{4}{25}} \left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2}\nn\\
%	      &= \sqrt{\frac{13}{10}}\,\pi^{-2/3}\,\f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6}\left[\int_0^{f_\text{ISCO}} df\, \f{f^{-7/3}}{S_\text{noise}(f)}\right]^{1/2} \label{eq:SNR_total_3IFOs_v2}.
\end{align}
%
%%
with $\rho_h, \rho_m,\rho_l$ given by Eq.~(\ref{eq:SNR_i}) and Table~\ref{table:network_params}.
%Although we made several simplifying assumptions to obtain $\rho_\text{tot}$, these should not affect our total SNR estimation too much.
%For example, averaging over all the angles underestimates the SNR of an interferometer with a near-ideal orientation with respect to the incoming GWs
%while also overestimating the SNR of a badly oriented detector. Rewriting Eq.~(\ref{eq:SNR_total}) as Eq.~(\ref{eq:SNR_total_3IFOs}) compensates
%for these.
%
%In Sec.~\ref{sec:GW170817} we test our model network with GW170817 whose signal got picked up by the LIGO detectors around $f=30$\,Hz.
%To do so we will change the lower limit of the integral in Eq.~(\ref{eq:SNR_total_3IFOs}) from $0$ to $30$.
%For our advance warning times we will further use separate initial frequencies for each inteferometer as we will explain in Sec.~\ref{sec:results}.

\subsection{Einstein Telescope configuration: triangular topology}\label{sec:ET_topo}
Current design of Einstein Telescope (ET) is based on a [equilateral] triangular configuration with 10\,km armlengths \cite{ET_doc}.
Within this equilateral triangle, ET will consist of three V-shaped cryogenic interferometers housed underground to
sigificantly reduce seismic and gravity-gradient noises \cite{ET_doc}.
This closed topology will allow the detector to form a null stream completely devoid of a signal \cite{Sathyaprakash:2012jk},
which can be used to rule out spurious events \cite{Wen:2005ui}.
As we will show below, the response function and the quality factor of the triangular configuration are independent of the azimuthal angle $\phi$
hence ET will have no blind spots \cite{Regimbau:2012ir}.
However, the $60^\circ$ arm separation reduces the strain sensitivity of a V-shaped interferometer by a factor of $\sin 60^\circ=\sqrt{3}/2$ compared
to an L-shaped interferometer with the same armlength. So for a single V-IFO the antenna patterns of (\ref{eq:F_plus}, \ref{eq:F_cross}) become
%
\be
F^1_+(\theta,\phi,\psi)= \f{\sqrt{3}}{2} F_+(\theta,\phi,\psi), \qquad F^1_\times(\theta,\phi,\psi)= \f{\sqrt{3}}{2} F_\times(\theta,\phi,\psi)\label{eq:F1plus_F1cross}\, ,
\ee
%
where the superscript $1$ labels one of the three V's of the triangle.
Since the other two V-IFOs lie in the same plane as the first one, their antenna patterns can be obtained simply by rotating in the azimuthal
direction by $120^\circ, 240^\circ$, respectively
%
\begin{align}
 F^2_{+,\times}(\theta,\phi,\psi) &= F^1_{+,\times}(\theta,\phi+2\pi/3,\psi), \label{eq:F2plus_F2cross}\\
 F^3_{+,\times}(\theta,\phi,\psi) &= F^1_{+,\times}(\theta,\phi-2\pi/3,\psi) \label{eq:F3plus_F3cross}.
\end{align}
%
The invidual antenna responses combine to cancel the $\phi$ and $\psi$ dependence of the combined power pattern
%
\be
F^2_\bigtriangleup \equiv \sum_{A=1}^3 \left(F^A_+\right)^2+\left(F^A_\times\right)^2 = \f{9}{32}\left(1+6\cos^2\theta + \cos^4\theta\right). \label{eq:ET_power_pattern}
\ee
%
From this, we immediately have that the RMS average is given by $\langle F^2_\bigtriangleup\rangle^{1/2} =3/\sqrt{10}$ and the minium value is $F_\bigtriangleup(\pi/2)=3/\sqrt{32}$;
%As $F^2 > 0$ for all $\theta$ with its RMS average $\sqrt{2/5}\simeq 63.2\%$ of optimal $F$ and its
thus a triangular detector made-up of three V-IFOs has all-sky coverage.

For the total SNR accumulated in a triangular detector we have
%
\begin{align}
\rho^2_{\text{tot},\bigtriangleup} & = \sum_{A=1}^3 \rho_A^2 \nn \\ % \f{c^2}{D^2}\left(\f{G M_c}{c^3}\right)^{5/3}
 & =  \f{5}{6}\, \pi^{-4/3} h_0^2\,|Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 \int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \, ,\label{eq:ET_SNRsq}
\end{align}
where
%
\be
|Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 = \sum_{A=1}^3 \left(F^A_+\right)^2\f{1+\cos^2\iota}{2}  + \left(F^A_\times\right)^2 \cos\iota \label{eq:Q_sq_ET}
\ee
%
which is independent of $\phi$ \cite{Regimbau:2012ir}. We once again take the RMS average of this quantity
%$Q(\theta,\phi,\psi,\iota)$ is the quality factor of the detector.
%
\be
\bar{Q}^2_\bigtriangleup\equiv\langle |Q(\theta,\phi,\psi,\iota)|_\bigtriangleup^2 \rangle_{\theta,\phi,\psi,\iota} = \f{9}{25}\label{eq:RMS_ET}
\ee
%
which is larger than a single L-shaped IFO's RMS average of 4/25 given in Eq.~(\ref{eq:RMS_th_ph_ps_iota}).

As we can not a priori know the source sky position or orientation we will model the response of ET using the RMS-averaged SNR %(\ref{eq:ET_SNRsq})
%
\be
\rho_{\text{tot},\bigtriangleup} = \f{6}{5}A\, h_0 \left[\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)}\right]^{1/2} \label{eq:ET_SNR},
\ee
where $h_0$ is given by Eq.~(\ref{eq:h0}) and $A=\pi^{-2/3}\sqrt{5/24}$ as before. % from Sec.\ref{sec:LIGO_topo}. 
Accordingly, we define an RMS-averaged frequency-domain strain
%
\begin{align}
\tilde{H}_\bigtriangleup(f)&\equiv A\,  \bar{Q}_\bigtriangleup\, h_0 f^{-7/6} \nn \\
   & = \f{1}{2}\sqrt{\f{3}{10}}\, \f{c}{D}\left(\f{G M_c}{c^3}\right)^{5/6} f^{-7/6}. \label{eq:H_delta}
\end{align}
%
This is ET's sky-averaged response to a BNS inspiral at luminosity distance $D$ and chirp mass $M_c$.
%It is this strain that we use as a typical ET BNS source; 
Thus, $\tilde{H}_\bigtriangleup(f)$ is like the $\tilde{H}_i(f)$ of Table~\ref{table:network_params},
but different in that  $\tilde{H}_\bigtriangleup(f)$ represents the total response of the entire three V-IFO network whereas $\tilde{H}_i(f)$
is the response of the $i^\text{th}$ L-IFO.
%whereas in the case of a three L-IFO network we represented the detector response due to a single source in terms of three different
%values $\tilde{H}_i(f)$, for ET we will only consider $\tilde{H}_\bigtriangleup(f)$ as we have already angle-averaged over all three V-IFOs antenna patterns.
%$\tilde{H}_\bigtriangleup(f)$, as given by Eq.~(\ref{eq:H_delta}), depends only on the luminosity distance and the chirp mass.

In Sec.~\ref{sec:results} we will consider BNS systems inspiralling at varying luminosity distances as typical sources for the ALV and ET networks.
To compute the resulting SNRs, we will slightly modify the integrals in Eqs.~(\ref{eq:SNR_i}, \ref{eq:ET_SNR}) and 
use design sensitivities of A-LIGO and ET's B and C configurations for $\sqrt{S_n(f)}$.
Before we present our results, we list the various idealizations that we have adopted to simplify our computations.
